---
title: "p8105_hw6_rac2224"
author: "Rachel Crowley"
date: "11/24/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "90%"
)

options(ggplot2.continuous.colour = "viridis",
        ggplot2.continuous.fill = "viridis"
        )

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))

set.seed(1)
```

## Problem 1

Loading and cleaning the data for regression analysis 

```{r}
bwt_data = read_csv("data/birthweight.csv") %>%
  rename(
    baby_sex = babysex,
    baby_head = bhead,
    baby_length = blength,
    birthweight = bwt,
    delivery_weight = delwt,
    family_income = fincome,
    father_race = frace,
    gestational_age = gaweeks,
    malformations = malform,
    menarche_age = menarche,
    mother_height = mheight,
    mother_age = momage, 
    mother_race = mrace,
    previous_low_birthweight = pnumlbw,
    prior_small_for_gestational = pnumsga, 
    pre_pregnancy_bmi = ppbmi,
    pre_pregnancy_weight = ppwt,
    smoking_during_pregnancy = smoken, 
    pregnancy_weight_gain = wtgain
  ) %>%
  mutate(
    baby_sex = as.factor(baby_sex),
    father_race = as.factor(father_race),
    malformations = as.factor(malformations),
    mother_race = as.factor(mother_race) )%>%
  mutate(
    baby_sex = recode(baby_sex, '1' = 'Male', '2' = 'Female'),
    father_race = recode(father_race, '1' = 'White', '2' = 'Black', '3' = 'Asian', '4' = 'Puerto Rican', '8' = 'Other', '9' = 'Unknown'),
    malformations = recode(malformations, '0' = 'Absent', '1' = 'Present'),
    mother_race = recode(mother_race, '1' = 'White', '2' = 'Black', '3' = 'Asian', '4' = 'Puerto Rican', '8' = 'Other')
  ) %>%
  mutate(
    baby_sex = fct_infreq(baby_sex),
    father_race = fct_infreq(father_race),
    malformations = fct_infreq(malformations),
    mother_race = fct_infreq(mother_race)
  )
 
colSums(is.na(bwt_data))

```

* baby_sex, mother_race, father_race, and malformations were converted from numeric to factor variables. There were no missing values across all variables in the dataset. 

Building regression models 

* For the proposed regression model for birthweight, the following variables were included in the model as predictors, based on existing literature that support an associaton between these factors and birthweight:  average number of cigarettes smoked per day during pregnancy (smoking_during_pregnancy), mother's age at delivery in years (mother_age), mother’s pre-pregnancy BMI (pre_pregnancy_bmi), and gestational age in weeks (gestational_age). It is hypothesized that for a one unit increase in the average number of cigarettes smoked per day during pregnancy, birthweight will decrease; for an increase in age at delivery in years, birthweight will decrease; for a one unit increase in mother's pre-pregnancy BMI, birthweight will increase; for an increase in gestational age in weeks, birthweight will increase. 

Fitting the model and plot of model residuals against fitted value

```{r}

fit = lm(birthweight ~ smoking_during_pregnancy + mother_age + pre_pregnancy_bmi + gestational_age, data = bwt_data)

summary(fit)$coef

fit %>% 
  broom::tidy() %>%
   knitr::kable(digits = 3)

bwt_data %>%
  modelr::add_residuals(fit) %>%
  modelr::add_predictions(fit) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point() + 
  geom_hline(aes(yintercept = 0, color = "red")) + 
  labs(
    title = "Model residuals against fitted values for proposed regression model for birthweight",
    y = "Residuals",
    x = "Fitted Values",
    caption = "Data from Birthweight.csv"
  )
```
 
Fitting two other models 
 
```{r}
fit2 = lm(birthweight ~ baby_length + gestational_age, data = bwt_data)

summary(fit2)$coef

fit2 %>% 
  broom::tidy() %>%
   knitr::kable(digits = 3)

bwt_data %>%
  modelr::add_residuals(fit2) %>%
  modelr::add_predictions(fit2) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point() + 
  geom_hline(aes(yintercept = 0, color = "red")) + 
  labs(
    title = "Model residuals against fitted values for regression model with baby length and gestational age",
    y = "Residuals",
    x = "Fitted Values",
    caption = "Data from Birthweight.csv"
  )

fit3 = lm(birthweight ~ baby_head*baby_length*baby_sex, data = bwt_data)


summary(fit3)$coef

fit3 %>% 
  broom::tidy() %>%
   knitr::kable(digits = 3)

bwt_data %>%
  modelr::add_residuals(fit3) %>%
  modelr::add_predictions(fit3) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point() + 
  geom_hline(aes(yintercept = 0, color = "red")) + 
  labs(
    title = "Model residuals against fitted values for regression model with head size, length, and sex interaction",
    y = "Residuals",
    x = "Fitted Values",
    caption = "Data from Birthweight.csv"
  )

```

* describe plots and process for model building 

Comparing regression models in terms of the cross-validated prediction error

```{r, warning = F}

cv_df = 
  crossv_mc(bwt_data, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble) 
  ) %>%
  mutate(
    linear_mod1  = map(train, ~lm(birthweight ~ smoking_during_pregnancy + mother_age + pre_pregnancy_bmi + gestational_age,     data = bwt_data)),
    linear_mod2  = map(train, ~lm(birthweight ~ baby_length + gestational_age, data = bwt_data)),
    linear_mod3  = map(train, ~lm(birthweight ~ baby_head*baby_length*baby_sex, data = bwt_data))) %>% 
  mutate(
    rmse_linear1 = map2_dbl(linear_mod1, test, ~rmse(model = .x, data = .y)),
    rmse_linear2 = map2_dbl(linear_mod2, test, ~rmse(model = .x, data = .y)),
    rmse_linear3 = map2_dbl(linear_mod3, test, ~rmse(model = .x, data = .y))) %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse", 
    names_prefix = "rmse_") %>%
      ggplot(aes(x = model, y = rmse)) + geom_violin() + labs(
        title = "Root mean squared error distribution by model type",
        x = "Model",
        y = "Root Mean Squared Error: RMSE",
        caption = "Data from birthweight.csv"
      )
  
cv_df
```